---
title: "Correct the Cloze probabilities"
author: "Miriam Schulz"
date: "2025-02-25"
output:
  html_document:
    number_sections: true
    toc: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r libraries, message=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
rm(list = ls())
```


# About

This script checks the Stone et al. 2023 target cloze values with respect to the Nicenboim et al. 2020 Cloze data,
and corrects them, removing items with zero cloze targets.


# Calculate Cloze manually

Read the stimuli, filter for the relevant conditions and check the number of items:

```{r read-data}
# Read in the data
dat <- read.csv("answers_annotated_before_noun_CORRECTED.tsv", header=TRUE, sep = "\t")
stim <- read.csv("stimuli_stone_preprocessed_surprisal.csv", header=TRUE)
```

```{r}
# Manually compute cloze probabilities
cloze <- dat %>%
  dplyr::group_by(item, condition) %>%  # Group by item and condition first
  dplyr::mutate(total_responses = n()) %>%  # Count total responses per item-condition
  dplyr::group_by(item, condition, stemnoun) %>%  # Then group by response
  dplyr::summarise(count = n(), total_responses = first(total_responses), .groups = "drop") %>%
  dplyr::mutate(cloze_probability = count / total_responses)

head(cloze)

cloze <- cloze %>% 
  dplyr::filter(condition == "Constraining_a") %>% 
  dplyr::rename(clozenoun = stemnoun)

# Ad the cloze probabilities to the stimuli
dat_cloze <- merge(stim, cloze,
                   by = c("item"),
                   all.x = TRUE,
                   all.y = FALSE)

dat_cloze <- dat_cloze %>% 
  arrange(item, cond_label)

# Keep only those cloze responses that match a target in the items,
# i.e., discard non-target cloze responses;
# however, this will also remove all zero cloze B conditions!
dat_cloze <- dat_cloze %>%
  filter(stemnoun == clozenoun | noun == clozenoun)
length(unique(dat_cloze$item))  # some items missing

# summary(as.factor(dat_cloze$noun == dat_cloze$clozenoun))
# summary(as.factor(dat_cloze$stemnoun == dat_cloze$clozenoun))

# To recover the missing zero cloze b conditions, create a new data frame 
# from the stimuli containing only those.
# First, reduce the data frame to the columns from the stimuli file:
dat_cloze_reduced <- dat_cloze %>% 
  select(colnames(stim))
# Then, get the rows that are in the stimuli but not in the cloze annotated
# stimuli:
stim_missing <- dplyr::anti_join(stim, dat_cloze_reduced,
                                 by = c("item", "cond_label"))

# Finally, manually fill in the columns from the cloze data set, 
# like zero cloze probability:
setdiff(colnames(dat_cloze), colnames(stim_missing))
stim_missing$condition <- "Constraining_a"
stim_missing$clozenoun <- "NA"
stim_missing$count <- 0
stim_missing$total_responses <- "NA"
stim_missing$cloze_probability <- 0.0

# Merge
dat_cloze_full <- rbind(dat_cloze, stim_missing)
dat_cloze_full <- unique(dat_cloze_full)
length(unique(dat_cloze_full$item))  # all items are there now
xtabs(~ cond_label, dat_cloze_full)  # this is concerning...
xtabs(~ item + cond_label, dat_cloze_full)  # 2x 71b, 2x 132a and 133a
dat_duplicates <- dat_cloze_full %>% 
  filter(item %in% c(71, 132, 133)) %>% 
  arrange(item, cond_label)
# write.csv(dat_duplicates, "delete.csv")

# There is an issue with 3 items, the clozenoun contains different labels.
# Manually delete the duplicates for these items:
dat_cloze_full <- dat_cloze_full %>% 
  filter(!(item == 71 & clozenoun == "Verkäuferin")) %>% 
  filter(!(item == 132 & clozenoun == "Fotografen")) %>% 
  filter(!(item == 133 & clozenoun == "Briefmarke"))

range(xtabs(~ item + cond_label, dat_cloze_full))
xtabs(~ cond_label, dat_cloze_full)
```


# Check problematic items

Check items with problematic Cloze values and inspect them manually (looking at the raw Cloze responses in `answers_annotated_before_noun.tsv`; corrections are saved in `answers_annotated_before_noun_CORRECTED.tsv`).

```{r}
dat_impossible <- dat_cloze_full %>% 
  filter(cond_label == "a" & cloze_probability < 0.2)

dat_cloze_full$cloze_probability <- round(dat_cloze_full$cloze_probability, 2)
summary(dat_cloze_full$cloze.target == dat_cloze_full$cloze_probability)
dat_inspect <- dat_cloze_full %>% 
  filter(cloze.target != cloze_probability) %>% 
  select(item, cond_label,
         noun, stemnoun, clozenoun,
         cloze.target, cloze.target_smth, cloze_probability,
         count, total_responses) %>% 
  arrange(item, cond_label)
```

Some items seem to have made-up cloze probabilities, OR their cloze probabilities were derived from other targets than the ones used in a. 

Delete these items:

```{r}
items_to_delete <- c(13, 76, 44, 28, 78, 21, 181, 276)

dat_cloze_full <- dat_cloze_full %>% 
  filter(!(item %in% items_to_delete))
```


# Stone et al. vs. corrected Cloze probabilities

Check cloze calculated by Stone vs. cloze as I calculated:

```{r}
# Stone cloze probs
dat_cloze_full %>%
  dplyr::group_by(cond_label) %>%
  dplyr::summarise(MeanCloze = mean(cloze.target),
                   ClozeSD = sd(cloze.target),
                   MinCloze = min(cloze.target),
                   MaxCloze = max(cloze.target))

# My cloze probs
dat_cloze_full %>%
  dplyr::group_by(cond_label) %>%
  dplyr::summarise(MeanCloze = mean(cloze_probability),
                   ClozeSD = sd(cloze_probability, na.rm = TRUE),
                   MinCloze = min(cloze_probability),
                   MaxCloze = max(cloze_probability))
```


# Remove zero cloze items

Check how many items have zero cloze:

```{r}
dat_zero_cloze <- dat_cloze_full %>% 
  filter(cloze_probability == 0)
summary(as.factor(dat_zero_cloze$cond_label))
nrow(dat_zero_cloze)

length(unique(dat_zero_cloze$item)) / length(unique(dat_cloze_full$item)) * 100
```

~27% of all items have a zero cloze target.

```{r}
ggplot(data=dat_cloze_full,
       aes(cloze_probability, fill = cond_label, color = cond_label)) + 
  geom_density(alpha = 0.6) + 
  scale_fill_manual(values = c("navy", "magenta3")) +
  scale_color_manual(values = c("navy", "magenta3"), guide = "none") +
  labs(title = "Cloze probabilities by condition",
       fill = "Condition")
```

Remove the zero cloze target items:

```{r}
dat_keep <- dat_cloze_full %>%
  filter(!(item %in% dat_zero_cloze$item))
length(unique(dat_keep$item))

dat_keep %>%
  dplyr::group_by(cond_label) %>%
  dplyr::summarise(MeanCloze = mean(cloze_probability),
                   ClozeSD = sd(cloze_probability, na.rm = TRUE),
                   MinCloze = min(cloze_probability),
                   MaxCloze = max(cloze_probability))

ggplot(data=dat_keep,
       aes(cloze_probability, fill = cond_label, color = cond_label)) + 
  geom_density(alpha = 0.6) + 
  scale_fill_manual(values = c("navy", "magenta3")) +
  scale_color_manual(values = c("navy", "magenta3"), guide = "none") +
  labs(title = "Cloze probabilities by condition",
       fill = "Condition")
```

# Export

Annotate which items to keep/discard due to cloze, and export.

```{r}
dat_cloze_full$keep_item <- ifelse(dat_cloze_full$item %in% dat_keep$item,
                                   "keep", "discard")
dat_cloze_full <- dat_cloze_full %>% 
  arrange(item, cond_label)
write.csv(dat_cloze_full, "nicenboim_stone_cloze_corrected.csv", row.names = FALSE)
```

