---
title: "Correct the Cloze probabilities"
author: "Miriam Schulz"
date: "2025-02-25"
output:
  html_document:
    number_sections: true
    toc: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r libraries, message=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
rm(list = ls())
```


# About

This script checks the Stone et al. 2023 target cloze values with respect to the Nicenboim et al. 2020 Cloze data,
and corrects them, removing items with zero cloze targets.


# Calculate Cloze manually

Read the stimuli, filter for the relevant conditions and check the number of items:

```{r read-data}
# Read in the data
dat <- read.csv("answers_annotated_before_noun_CORRECTED.tsv", header=TRUE, sep = "\t")
stim <- read.csv("stimuli_stone_preprocessed_surprisal.csv", header=TRUE)
```

Remove trailing white spaces that would prevent the words from being merged:

```{r trim-ws}
dat$stemnoun <- trimws(dat$stemnoun, which = "both")
dat$word <- trimws(dat$word, which = "both")
```

Compute word form and stem noun cloze probabilities:

```{r compute-cloze}
# Manually compute the stem noun cloze probabilities
cloze <- dat %>%
  dplyr::group_by(item, condition) %>%  # Group by item and condition first
  dplyr::mutate(total_responses_stemnoun = n()) %>%  # Count total responses per item-condition
  dplyr::group_by(item, condition, stemnoun) %>%  # Then group by response
  dplyr::summarise(count_stemnoun = n(), total_responses_stemnoun = first(total_responses_stemnoun), .groups = "drop") %>%
  dplyr::mutate(cloze_stemnoun = count_stemnoun / total_responses_stemnoun) %>% 
  filter(condition == "Constraining_a") %>% 
  arrange(item, -cloze_stemnoun)

# Export for individual analysis
write.csv(cloze, "nicenboim_cloze_allanswers_stemnoun.csv", row.names = FALSE)

# Manually compute the word form cloze probabilities
cloze_wordform <- dat %>%
  dplyr::group_by(item, condition) %>%  # Group by item and condition first
  dplyr::mutate(total_responses_wordform = n()) %>%  # Count total responses per item-condition
  dplyr::group_by(item, condition, word_CORRECTED, stemnoun) %>%  # Then group by response
  dplyr::filter(condition == "Constraining_a") %>% 
  dplyr::select(item, word_CORRECTED,	stemnoun,	total_responses_wordform) %>%   
  dplyr::rename(word = word_CORRECTED) %>% 
  dplyr::summarise(count_wordform = n(), total_responses_wordform = first(total_responses_wordform), .groups = "drop") %>%
  dplyr::mutate(cloze_wordform = count_wordform / total_responses_wordform) %>%
  arrange(item, -cloze_wordform)

# Merge
cloze_wordform <- merge(cloze, cloze_wordform,
                         by = c("item", "condition","stemnoun"),
                         all = TRUE)
cloze_wordform <- cloze_wordform %>%
  select(item, condition, stemnoun, word,
          cloze_stemnoun, cloze_wordform,
          count_stemnoun, total_responses_stemnoun,
          count_wordform, total_responses_wordform) %>% 
  arrange(item, -cloze_stemnoun, -cloze_wordform)

# Export for individual analysis
write.csv(cloze_wordform, "nicenboim_cloze_allanswers_wordform.csv",
          row.names = FALSE)
```

```{r merge-cloze-with-stimuli}
# Merge the word form cloze responses and their probabilities into a single cell
# from long to wide format, separated by "/", if there is more than one:
# cloze_wordform_wide <- cloze_wordform %>%
#   dplyr::mutate(cloze_stemnoun = round(cloze_stemnoun, 4),
#                 cloze_wordform = round(cloze_wordform, 4)) %>% 
#   dplyr::group_by(item, condition, stemnoun, cloze_stemnoun,
#                   total_responses_stemnoun) %>% 
#   dplyr::rename(clozenoun = stemnoun,
#                 n_cloze_responses = total_responses_stemnoun) %>% 
#   dplyr::summarise(
#     word = paste(word, collapse = "/"), 
#     cloze_wordform = paste(cloze_wordform, collapse = "/"),
#     .groups = "drop"
#   )

cloze_wordform_wide <- cloze_wordform %>%
  dplyr::mutate(
    cloze_stemnoun = round(cloze_stemnoun, 4),
    cloze_wordform = round(cloze_wordform, 4)
  ) %>%
  dplyr::group_by(item, condition, stemnoun, cloze_stemnoun,
                  total_responses_stemnoun) %>%
  dplyr::rename(
    clozenoun = stemnoun,
    n_cloze_responses = total_responses_stemnoun
  ) %>%
  dplyr::summarise(
    wordform_highest = first(word),
    cloze_wordform_highest = first(cloze_wordform),
    wordform_other = paste(word[-1], collapse = "/"),
    cloze_wordform_other = paste(cloze_wordform[-1], collapse = "/"),
    .groups = "drop"
  )

write.csv(cloze_wordform_wide, "nicenboim_cloze_allanswers_wordform_wide.csv",
          row.names = FALSE)

# cloze <- cloze %>%
#   dplyr::rename(clozenoun = stemnoun)

# Add the cloze probabilities to the stimuli
dat_cloze <- merge(stim, cloze_wordform_wide,
                   by = c("item"),
                   all.x = TRUE,
                   all.y = FALSE)

dat_cloze <- dat_cloze %>% 
  arrange(item, cond_label)

# Keep only those cloze responses that match a target in the items,
# i.e., discard non-target cloze responses;
# however, this will also remove all zero cloze B conditions!
dat_cloze <- dat_cloze %>%
  filter(stemnoun == clozenoun | noun == clozenoun)
length(unique(dat_cloze$item))  # some items missing

# summary(as.factor(dat_cloze$noun == dat_cloze$clozenoun))
# summary(as.factor(dat_cloze$stemnoun == dat_cloze$clozenoun))

# To recover the missing zero cloze b conditions, create a new data frame 
# from the stimuli containing only those.
# First, reduce the data frame to the columns from the stimuli file:
dat_cloze_reduced <- dat_cloze %>% 
  select(colnames(stim))
# Then, get the rows that are in the stimuli but not in the cloze annotated
# stimuli:
stim_missing <- dplyr::anti_join(stim, dat_cloze_reduced,
                                 by = c("item", "cond_label"))

# Finally, manually fill in the columns from the cloze data set, 
# like zero cloze probability:
setdiff(colnames(dat_cloze), colnames(stim_missing))
stim_missing$condition <- "Constraining_a"
stim_missing$clozenoun <- "NA"
stim_missing$cloze_stemnoun <- 0
stim_missing$n_cloze_responses <- "NA"
stim_missing$wordform_highest <- "NA"
stim_missing$cloze_wordform_highest <- 0.0
stim_missing$wordform_other <- "NA"
stim_missing$cloze_wordform_other <- 0.0
# stim_missing$word <- "NA"
# stim_missing$cloze_wordform <- 0.0
# summary(colnames(dat_cloze) == colnames(stim_missing))

# Merge
dat_cloze_full <- rbind(dat_cloze, stim_missing)
dat_cloze_full <- unique(dat_cloze_full)
```

```{r check-duplicates}
# Check that all items are there
length(unique(dat_cloze_full$item))  # all items are there now
xtabs(~ cond_label, dat_cloze_full)  # this is concerning...
#xtabs(~ item + cond_label, dat_cloze_full)  # previously 2x 71b, 2x 132a and 133a

# Check if/which items are there twice:
tab <- as.data.frame(xtabs(~ item + cond_label, dat_cloze_full))
row_sums <- tab %>%
  dplyr::group_by(item) %>%
  dplyr::summarise(total = sum(Freq)) %>%
  arrange(desc(total)) # Sort by decreasing frequency
# Reorder item factor levels based on sorted counts
tab$item <- factor(tab$item, levels = row_sums$item)
tab <- xtabs(Freq ~ item + cond_label, data = tab)
tab  # items 5a and 113a are there twice

dat_duplicates <- dat_cloze_full %>% 
  # filter(item %in% c(71, 132, 133)) %>%   # previous duplicate items
  filter(item %in% c(5, 113)) %>%  # new duplicate items
  arrange(item, cond_label)
# write.csv(dat_duplicates, "delete.csv")

# There is an issue with 3 items, the clozenoun contains different labels.
# Manually delete the duplicates for these items:
# dat_cloze_full <- dat_cloze_full %>% 
#   filter(!(item == 71 & clozenoun == "Verkäuferin")) %>% 
#   filter(!(item == 132 & clozenoun == "Fotografen")) %>% 
#   filter(!(item == 133 & clozenoun == "Briefmarke"))
dat_cloze_full <- dat_cloze_full %>% 
  filter(!(item == 5 & clozenoun == "Eisen")) %>% 
  filter(!(item == 113 & clozenoun == "Buch"))

# Check that this fixed any imbalances:
range(xtabs(~ item + cond_label, dat_cloze_full))
xtabs(~ cond_label, dat_cloze_full)
```


# Check problematic items

Check items with problematic Cloze values and inspect them manually (looking at the raw Cloze responses in `answers_annotated_before_noun.tsv`; corrections are saved in `answers_annotated_before_noun_CORRECTED.tsv`).

```{r check-items}
dat_impossible <- dat_cloze_full %>% 
  filter(cond_label == "a" & cloze_stemnoun < 0.2)

# Manually inspect items for which my calculated stem noun cloze probability
# does not match the cloze probability calculated by Stone et al.: 
dat_cloze_full$cloze_stemnoun <- round(dat_cloze_full$cloze_stemnoun, 2)
summary(dat_cloze_full$cloze.target == dat_cloze_full$cloze_stemnoun)
dat_inspect <- dat_cloze_full %>% 
  filter(cloze.target != cloze_stemnoun) %>% 
  select(item, cond_label,
         noun, stemnoun, clozenoun,
         cloze.target, cloze.target_smth, cloze_stemnoun,
         n_cloze_responses) %>% 
  arrange(item, cond_label)
unique(dat_inspect$item)
```

Manual inspection reveals that some items seem to have altogether made-up cloze probabilities, OR their cloze probabilities were derived from other targets than the ones used in a. 

Delete these items:

```{r define-items-to-delete}
items_to_delete <- c(13, 76, 44, 28, 78, 21, 181, 276)

# dat_cloze_full <- dat_cloze_full %>% 
#   filter(!(item %in% items_to_delete))
```


# Stone et al. vs. corrected Cloze probabilities

Check cloze calculated by Stone vs. cloze as I calculated:

```{r check-cloze-probs}
# Stone cloze probs
dat_cloze_full %>%
  dplyr::group_by(cond_label) %>%
  dplyr::summarise(MeanCloze = mean(cloze.target),
                   ClozeSD = sd(cloze.target),
                   MinCloze = min(cloze.target),
                   MaxCloze = max(cloze.target))

# My cloze probs
dat_cloze_full %>%
  dplyr::group_by(cond_label) %>%
  dplyr::summarise(MeanCloze = mean(cloze_stemnoun),
                   ClozeSD = sd(cloze_stemnoun, na.rm = TRUE),
                   MinCloze = min(cloze_stemnoun),
                   MaxCloze = max(cloze_stemnoun))
```


# Remove zero cloze items

Check how many items have zero cloze:

```{r check-zero-cloze}
dat_zero_cloze <- dat_cloze_full %>% 
  filter(cloze_stemnoun == 0)
nrow(dat_zero_cloze)
summary(as.factor(dat_zero_cloze$cond_label))

length(unique(dat_zero_cloze$item)) / length(unique(dat_cloze_full$item)) * 100
```

~30% of all items have a zero cloze target.  
(Prior to correcting more cloze values, this was ~27% of all items.)

```{r plot-cloze}
ggplot(data=dat_cloze_full,
       aes(cloze_stemnoun, fill = cond_label, color = cond_label)) + 
  geom_density(alpha = 0.6) + 
  scale_fill_manual(values = c("navy", "magenta3")) +
  scale_color_manual(values = c("navy", "magenta3"), guide = "none") +
  labs(title = "Cloze probabilities by condition",
       fill = "Condition")
```

Remove the zero cloze target items:

```{r remove-zero-cloze}
dat_keep <- dat_cloze_full %>%
  filter(!(item %in% dat_zero_cloze$item) &
           !(dat_cloze_full$item %in% items_to_delete))
length(unique(dat_keep$item))

dat_keep %>%
  dplyr::group_by(cond_label) %>%
  dplyr::summarise(MeanCloze = mean(cloze_stemnoun),
                   ClozeSD = sd(cloze_stemnoun, na.rm = TRUE),
                   MinCloze = min(cloze_stemnoun),
                   MaxCloze = max(cloze_stemnoun))

ggplot(data=dat_keep,
       aes(cloze_stemnoun, fill = cond_label, color = cond_label)) + 
  geom_density(alpha = 0.6) + 
  scale_fill_manual(values = c("navy", "magenta3")) +
  scale_color_manual(values = c("navy", "magenta3"), guide = "none") +
  labs(title = "Cloze probabilities by condition",
       fill = "Condition")
```

# Export

Annotate which items to keep/discard due to cloze, and export.

```{r export-corrected}
dat_cloze_full$keep_item <- ifelse(dat_cloze_full$item %in% dat_keep$item,
                                   "keep", "discard")
dat_cloze_full <- dat_cloze_full %>% 
  arrange(item, cond_label)
write.csv(dat_cloze_full, "nicenboim_stone_cloze_corrected.csv", row.names = FALSE)
```
