---
title: "Check + fix the excluded items"
author: "Miriam Schulz"
date: "2025-02-25"
output:
  html_document:
    number_sections: true
    toc: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r libraries, message=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
rm(list = ls())
```


# About

This script checks the items that were removed from the Stone 2023 stimuli in an analyis of the cloze probabilities using `correct_cloze.Rmd`.

It might be possible to use them with different targets.

The items to check are:

- 8 items were excluded from the Stone et al. 2023 stimuli because they (prorbably mistakenly) did not use the highest cloze target for condition a,.
- 58 items were excluded from the Stone et al. 2023 stimuli because their condition b targets had zero cloze probability.


# Read data and check manually

```{r}
dat <- read.csv("nicenboim_stone_cloze_corrected.csv", header = TRUE)
cloze <- read.csv("nicenboim_cloze_allanswers.csv", header = TRUE)
cloze_all <- read.csv("answers_annotated_before_noun_CORRECTED.tsv",
                      sep = "\t", header = TRUE)
```


# 8 items with wrong condition a targets

## Check

```{r}
critical_items <- c(13, 21, 28, 44, 76, 78, 181, 276)

dat_wrongA <- dat %>% 
  filter(item %in% critical_items)
cloze_wrongA <- cloze %>% 
  filter(item %in% critical_items,
         condition == "Constraining_a") %>% 
  arrange(item, -cloze_probability)
cloze_all_wrongA <- cloze_all %>% 
  filter(item %in% critical_items,
         condition == "Constraining_a")
```

Manually inspect the sentence contexts in `dat_wrongA` and the potential targets in `cloze_wrongA` (as well as the non-truncated forms in the original cloze data, `cloze_all_wrongA`) to choose new targets.


## Choose new targets

New manually chosen targets:

- 13: a: Krankenwagen (stemnoun: Wagen), b: Notdienst / Kollegen
- 21: a: Update, b: Programm
- 28: no solution (only Tasche vs. Tüte)
- 44: a: Bleistift, b: Anspitzer
- 76: a: Stelle, b: Gasse / Bank
- 78: a: Schaufel, b: Schippe
- 181: a: Nachricht b: Mitteilung (BUT: SYNONYMS :/)
- 276: no solution (low Cloze targets only)



# 58 items with zero cloze condition b targets

## Check

```{r}
dat <- dat %>%
  filter(keep_item == "discard" & !(item %in% critical_items))
cloze <- cloze %>% 
  filter(item %in% dat$item,
         condition == "Constraining_a") %>% 
  arrange(item, -cloze_probability)
cloze_all <- cloze_all %>% 
  filter(item %in% dat$item,
         condition == "Constraining_a")

# Add sentence context to cloze for faster selection:
cloze <- merge(unique(dplyr::select(cloze_all, item, sentence)), cloze,
               by = "item",
               all.x = FALSE, all.y = TRUE)

# Write to file for easier inspection
write.csv(cloze, "items_to_check_manually.csv", row.names = FALSE)
```

Manually inspect the data to choose new condition b targets.


## Choose new targets

New manually chosen targets (item - a target - b target; ? means unsure, - means no solution found):

- 24 Turm Wall
- 30 Visum Gesetz
- 46 - 
- 49 Laken Kostüm
- 50 Pumpe Yacht ?
- 53 - 
- 54 Regal Paket
- 60 Fernglas Teleskop
- 65 -
- 68 -
- 70 -
- 82 -
- 88 (Müll)eimer Korb
- 106 Stethoskop Gerät
- 109 -
- 115 Besen Schuh ?
- 126 Feuer Lager
- 132 -
- 138 Bürste Zigarette
- 144 Suppe Tasse
- 146 -
- 152 -
- 158 Karte Unterlage ?
- 161 -
- 163 Rede Geschichte / Danksagung [CORRECT CLOZE]
- 164 Gitter Netz ?
- 166 -
- 169 -
- 175 Taschentuch Küsschen ?
- 177 - 
- 188 - 
- 189 Teleskop Fernrohr ?
- 194 -
- 196 - 
- 201 -
- 204 - 
- 206 -
- 212 Uhr Alternative ?
- 213 Schürze Serviette
- 222 -
- 225 Witz Aufsatz
- 229 Kalender [CORRECT CLOZE] Zettel/Stift
- 232 -
- 233 Zelt Tarp ?
- 234 Untersetzer Lappen
- 235 Toaster Paninimaker ??
- 237 -
- 241 -
- 244 -
- 246 Pause Tasse
- 247 -
- 254 -
- 268 -
- 269 -
- 270 -
- 271 -
- 273 Stein Splitter ?
- 278 -

Summary:

- 16 items saved
- 10 items maybe
- 32 items discarded

Plus an additional 6 items potentially saved for the wrong a conditions, this yields:  
out of 58 + 8 discarded items:  

- 22 may be saved (but these include some synonym targets)
- 10 maybe
- 34 discarded