---
title: "Change targets"
author: "Miriam Schulz"
date: "2025-03-04"
output:
  html_document:
    number_sections: true
    toc: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r libraries, message=FALSE}
library("tidyverse")
library("ggplot2")
library("Rmisc")
library(stringr)
library(rlang)
library(corrplot)
rm(list = ls())
```


# About

This script changes targets in the Stone et al. 2023 experimental stimuli.

- Out of the 58 items that were removed because the condition b target used by Stone et al. 2023 was not attested in the Nicenboim et al. 2020 cloze responses, some items can be kept with different attested condition b target manually chosen from the Nicenboim et al. cloze data.
- Out of the 8 items that were removed because the condition a target chosen by Stone et al. 2023 was not the modal cloze response (probably copy-paste errors), 5 items can be kept by changing the condition a target to the modal response (however, some did not have a modal response).

# Read in data

```{r read-selected-stimuli}
dat <- read.csv("nicenboim_stone_items_MANUALLY_REPLACED_TARGETS.csv", header=TRUE)
```


# Inspect

Load the manually annotated items and inspect manually:

```{r separate-items-and-inspect}
keep <- dat %>% 
  filter(keep_item == "keep") %>% 
  arrange(keep_item, item, cond_label)
summary(keep$noun == keep$wordform_highest)
cat(length(unique(keep$item)), "items could be kept as they are.")

fixed <- dat %>% 
  filter(keep_item == "fixed") %>% 
  arrange(keep_item, item, cond_label)
summary(fixed$noun == fixed$wordform_highest)
cat(length(unique(fixed$item)), "items could additionally be fixed.")
cat(nrow(filter(fixed, !is.na(noun_new))), "nouns were replaced to fix the stimuli (in condition a and/or b).")

synonyms <- dat %>% 
  filter(keep_item == "synonym")
length(unique(synonyms$item))

discard <- dat %>% 
  filter(keep_item == "discard") 
cat(length(unique(discard$item)), "items could not be fixed.")

inspect <- dat %>% 
  filter(!(item %in% keep$item) & !(item %in% discard$item)) %>% 
  arrange(keep_item, item, cond_label)
cat(length(unique(inspect$item)), "items have other issues and need to be manually inspected.")
unique(inspect$keep_item)

write.csv(inspect, "critical_items_to_inspect_manually.csv", row.names = FALSE)
```

Examine the `inspect` items manually.


# Replace targets

Change the target nouns and stems and export:

```{r merge-keep-df}
# Replace the targets
dat_changed <- dat
dat_changed$noun_stone <- dat_changed$noun
# dat_changed$stemnoun_stone <- dat_changed$stemnoun
dat_changed <- dat_changed %>% 
  mutate(noun = if_else(noun_new != "", noun_new, noun),
         noun_changed = if_else(noun_new != "", "yes", ""))
dat_changed <- dat_changed %>% 
  mutate(stemnoun = if_else(stemnoun_new != "", stemnoun_new, stemnoun),
         stemnoun_changed = if_else(stemnoun_new != "", "yes", ""))
dat_changed <- dat_changed %>% 
  filter(final_decision != "discard")

n_changed_nouns <- dat_changed %>% 
  filter(noun_changed == "yes") %>% 
  unique() %>% 
  nrow()
n_changed_nouns_items <- dat_changed %>% 
  filter(noun_changed == "yes") %>% 
  select(item) %>%
  unique() %>% 
  nrow()
cat(n_changed_nouns, "total word forms were changed in the items, concerning",
    n_changed_nouns_items, "items.")

n_changed_stems <- dat_changed %>% 
  filter(stemnoun_changed == "yes") %>% 
  unique() %>% 
  nrow()
n_changed_stems_items <- dat_changed %>% 
  filter(stemnoun_changed == "yes") %>% 
  select(item) %>%
  unique() %>% 
  nrow()
cat(n_changed_stems, "total stems were changed in the items, concerning",
    n_changed_stems_items, "items.")
```


# Update the Cloze values

## Update stem and highest word form Cloze

Update the cloze values for the changed targets:

```{r update-cloze}
cloze_stemnoun <- read.csv("../nicenboim_2020_cloze/nicenboim_cloze_allanswers_stemnoun.csv", header = TRUE)
cloze_wordform <- read.csv("../nicenboim_2020_cloze/nicenboim_cloze_allanswers_wordform_wide.csv", header = TRUE)

# Update stem noun cloze
dat_changed$cloze_stemnoun_WRONG <- dat_changed$cloze_stemnoun
summary(as.factor(dat_changed$cloze_stemnoun_WRONG))
dat_changed$cloze_stemnoun <- NULL
dat_changed <- merge(dat_changed, cloze_stemnoun, all.x = TRUE)
dat_changed <- dat_changed %>% 
  arrange(item, cond_label)
summary(round(dat_changed$cloze_stemnoun, 2) == dat_changed$cloze_stemnoun_WRONG)

# Update highest word form
dat_changed$wordform_highest_WRONG <- dat_changed$wordform_highest
dat_changed$cloze_wordform_highest_WRONG <- dat_changed$cloze_wordform_highest
dat_changed$wordform_other_WRONG <- dat_changed$wordform_other
dat_changed$cloze_wordform_other_WRONG <- dat_changed$cloze_wordform_other
dat_changed$wordform_highest <- NULL
dat_changed$cloze_wordform_highest <- NULL
dat_changed$wordform_other <- NULL
dat_changed$cloze_wordform_other <- NULL
cloze_wordform <- cloze_wordform %>% 
  select(item, clozenoun,
         wordform_highest, cloze_wordform_highest,
         wordform_other, cloze_wordform_other) %>% 
  dplyr::rename(stemnoun = clozenoun)
dat_changed <- merge(dat_changed, cloze_wordform,
                   by = c("item", "stemnoun"),
                   all.x = TRUE,
                   all.y = FALSE)
dat_changed <- dat_changed %>% 
  arrange(item, cond_label)
```

```{r manual-inspection-changed-targets}
inspect_manually <- dat_changed %>% 
  select(item, cond_label,
         noun,
         stemnoun,
         wordform_highest, cloze_wordform_highest,
         wordform_other, cloze_wordform_other,
         noun_stone, stemnoun_stone,
         cloze_stemnoun, cloze_stemnoun_WRONG)
inspect_manually$equal <- inspect_manually$stemnoun == inspect_manually$stemnoun_stone
inspect_manually$equal_cloze <- round(inspect_manually$cloze_stemnoun, 2) == inspect_manually$cloze_stemnoun_WRONG

inspect_manually$noun_equal_highest_wordform <- inspect_manually$noun == inspect_manually$wordform_highest
inspect_manually$noun_equal_stemnoun <- inspect_manually$noun == inspect_manually$stemnoun
problems <- inspect_manually %>% 
  filter(noun_equal_highest_wordform == FALSE & noun_equal_stemnoun == FALSE)
```

There is only one remaining seemingly problematic item: for item  64a, "Putztuch" was chosen as the target because "Brillenputztuch" + "Putztuch" together have a higher Cloze probability (0.1818 + 0.1818 = 0.3636) than the stem "Tuch" alone (0.2727).

## Add chosen word form Cloze

```{r add-wordform-cloze}
cloze_wordform <- read.csv("../nicenboim_2020_cloze/nicenboim_cloze_allanswers_wordform.csv", header = TRUE)

# Update highest word form
cloze_wordform <- cloze_wordform %>% 
  select(item, word, cloze_wordform) %>% 
  dplyr::rename(noun = word)
dat_changed <- merge(dat_changed, cloze_wordform,
                   by = c("item", "noun"),
                   all.x = TRUE,
                   all.y = FALSE)
dat_changed <- dat_changed %>% 
  arrange(item, cond_label)

dat_changed %>%
  filter(is.na(cloze_wordform)) %>% 
  select(item, cond_label, noun, stemnoun)

# Set unattested word form cloze to zero
dat_changed$cloze_wordform <- ifelse(is.na(dat_changed$cloze_wordform), 0,
                                     dat_changed$cloze_wordform)
```

For the above items, the noun used was not attested as a word form (but instead included in a compound, for example, "Polster" occurred in "Stuhlpolster").

## Check

### Cloze values

Check stem Cloze by condition: 

```{r check-cloze-stems}
low_cloze_a <- dat_changed %>% 
  filter(cloze_stemnoun <= 0.2 & cond_label == "a") %>% 
  select(item, cloze_stemnoun, stemnoun) %>% 
  print()
high_cloze_b <- dat_changed %>% 
  filter(cloze_stemnoun > 0.2 & cond_label == "b") %>% 
  select(item, cloze_stemnoun, stemnoun) %>% 
  print()
```

```{r meancloze-stems}
summarySE(dat_changed,
          measurevar = "cloze_stemnoun",  
          groupvars = c("cond_label"),
          na.rm = TRUE)
```

Check word form Cloze by condition: 

```{r check-cloze-wordforms}
low_cloze_a <- dat_changed %>% 
  filter(cloze_wordform <= 0.2 & cond_label == "a") %>% 
  select(item, cloze_wordform, noun) %>% 
  print()
high_cloze_b <- dat_changed %>% 
  filter(cloze_wordform > 0.2 & cond_label == "b") %>% 
  select(item, cloze_wordform, noun) %>% 
  print()
```

```{r meancloze-wordforms}
summarySE(dat_changed,
          measurevar = "cloze_wordform",  
          groupvars = c("cond_label"),
          na.rm = TRUE)
```

### Comparisons

```{r compare-changes}
summary(dat_changed$stemnoun == dat_changed$stemnoun_stone)
summary(dat_changed$noun == dat_changed$noun_stone)
summary(dat_changed$noun == dat_changed$wordform_highest)

inspect_wordform_not_highest <- dat_changed %>% 
  filter(noun != wordform_highest) %>% 
  select(item, cond_label, stemnoun, cloze_stemnoun, noun, cloze_wordform,
         wordform_highest, cloze_wordform_highest,
         wordform_other, cloze_wordform_other)
```

```{r check-updates}
summary(dat_changed$cloze_stemnoun_WRONG == round(dat_changed$cloze_stemnoun, 2))
summary(dat_changed$wordform_highest_WRONG == dat_changed$wordform_highest)
summary(dat_changed$wordform_other_WRONG == dat_changed$wordform_other)
summary(dat_changed$cloze_wordform_other_WRONG == dat_changed$cloze_wordform_other)
```

```{r}
x <- dat %>% 
  filter(final_decision == "discard") %>% 
  select(item, reason_to_discard)
```


# Export

Rearrange the columns:

```{r rearrange-cols}
dat_changed_rearranged <- dat_changed %>% 
  select(item, cond_label,
         final_decision, keep_item,
         list_stone,
         context, sentence, pre_critical, determiner, adj,
         noun, noun_new_otheroptions, noun_stone, noun_changed,
         stemnoun, stemnoun_stone, stemnoun_changed,
         post_noun, continuation,
         question_stone, answer_stone,
         cloze_wordform, cloze_stemnoun, cloze_target_stone,
         wordform_highest, cloze_wordform_highest, 
         wordform_other, cloze_wordform_other,
         count_stemnoun, total_responses_stemnoun,
         cosine_targetContext, entropy, 
         fasttext_CHECK, surprisal_gerpt2_CHECK, surprisal_gerpt2_large_CHECK,
         surprisal_leo_7b_CHECK, surprisal_leo_13b_CHECK)

# Check which columns were eliminated
colnames(dat_changed)[!colnames(dat_changed) %in% colnames(dat_changed_rearranged)]
```

```{r export}
write.csv(dat_changed_rearranged, "stimuli_exp2_not_final.csv", row.names = FALSE)
write.csv(filter(dat_changed_rearranged, final_decision == "backup"), "backup_inspect.csv", row.names = FALSE)
```


# Check & remove duplicates

Nouns appearing more than 2-3x should be removed (a max. of 3x can be allowed, since all occurrences can then be placed in different lists).

Check the fixed items (with changed targets) for use of duplicate targets:

```{r subset-df}
keep_changed <- dat_changed %>% 
  filter(!(final_decision %in% c("discard", "backup")))
```

Stem nouns:

```{r check-duplicate-nouns-stems}
duplicates_stems <- keep_changed %>%
  dplyr::group_by(stemnoun) %>%
  dplyr::summarize(occurrences = n(),
            items = paste(unique(item), collapse = ", "),
            conditions = paste(unique(cond_label), collapse = ", ")) %>%
  dplyr::filter(occurrences > 1) %>% 
  arrange(-occurrences)

head(duplicates_stems, n = 20)
cat(nrow(duplicates_stems), "stem nouns are used more than once in the stimuli.")
unique(sort(duplicates_stems$stemnoun))

more_than_twice <- sum(duplicates_stems$occurrences[duplicates_stems$occurrences > 2])
cat(more_than_twice, "items contain stem nouns that are used more than twice.")

more_than_twice_keep_two <- sum(duplicates_stems$occurrences[duplicates_stems$occurrences > 2] - 2)
cat(more_than_twice_keep_two, "of these should be removed if two occurrences of each target can be kept.")
```

Word forms:

```{r check-duplicate-nouns-wordforms}
duplicates_wordforms <- keep_changed %>%
  dplyr::mutate(noun = gsub("[[:punct:]]", "", noun)) %>% 
  dplyr::group_by(noun) %>%
  dplyr::summarize(occurrences = n(),
            items = paste(unique(item), collapse = ", "),
            conditions = paste(unique(cond_label), collapse = ", ")) %>%
  dplyr::filter(occurrences > 1) %>% 
  arrange(-occurrences)

head(duplicates_wordforms, n = 20)
cat(nrow(duplicates_wordforms), "word forms are used more than once in the stimuli.")
unique(sort(duplicates_wordforms$noun))

more_than_twice <- sum(duplicates_wordforms$occurrences[duplicates_wordforms$occurrences > 2])
cat(more_than_twice, "items contain word forms that are used more than twice.")

more_than_twice_keep_two <- sum(duplicates_wordforms$occurrences[duplicates_wordforms$occurrences > 2] - 2)
cat(more_than_twice_keep_two, "of these should be removed if two occurrences of each target can be kept.")
```

Subset items for easy checking:

```{r}
check_items <- keep_changed %>% 
  # filter(item %in% c(117, 171, 182, 200)) %>%  # Hammer
  # filter(item %in% c(3, 148, 203)) %>%   # Anzug
  # filter(item %in% c(69, 250, 267)) %>%  # Brille
  # filter(item %in% c(74, 124, 198)) %>%  # Eimer
  # filter(item %in% c(66, 193, 209)) %>%  # Helm
  # filter(item %in% c(74, 137, 234)) %>%  # Lappen
  # filter(item %in% c(45, 176, 281)) %>%  # Messer
  # filter(item %in% c(39, 136, 184)) %>%  # Taschenrechner
  # filter(item %in% c(144, 151, 246)) %>%  # Tasse 
  filter(item %in% c(27, 64, 175, 211, 251)) %>%  # Tuch (STEM)
  # filter(item %in% c(1, 67, 114, 245)) %>%  # Schirm (STEM)
  select(item, cond_label, final_decision, reason_to_discard, keep_item,
         context, sentence, pre_critical, determiner, adj, noun, stemnoun)
```

Since the stems "Tuch" and "Schirm" appear in such different compound nouns, mostly with different word forms (Sonnenschirm-Regenschirm-Fallschirm-Schirm; Tuch-Putztuch-Taschentuch-Handtuch), up to 4-5 occurrences of each stem can be kept, as long as each identical word form is placed in a different list.

```{r check-n-remaining-items}
length(unique(keep_changed$item))
```

The remaining 158 items must be annotated for surprisal, word embeddings etc. and balanced for length and frequency by condition.