---
title: "Check the experimental stimuli"
author: "Miriam Schulz"
date: "2025-02-19"
output:
  html_document:
    number_sections: true
    toc: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r libraries, message=FALSE}
library("tidyverse")
library("ggplot2")
library("Rmisc")
library(stringr)
library(rlang)
library(corrplot)

rm(list = ls())
```


# About

This script checks the Stone et al. 2023 experimental stimuli and adds + analyzes metrics like similarity scores between the target words and target noun surprisal by 4 different LLMs.


# Inspect cloze probability per condition

Read the stimuli, filter for the relevant conditions and check the number of items:

```{r read-stimuli}
# Read in the data
# dat <- read.csv("experimental_stimuli.tsv", header=TRUE, sep = "\t")
dat <- read.csv("nicenboim_stone_cloze_corrected.csv", header=TRUE)

dat <- dat %>%
  filter(cond_label %in% c("a", "b"))

# How many items are there in total?
length(unique(dat$item))

# Remove items with zero cloze
dat_remove <- dat %>% 
  filter(keep_item == "discard")
dat <- dat %>% 
  filter(keep_item == "keep")
```


## Stem noun cloze

Check the distribution of Cloze probabilities:

Stone et al. "unsmoothed":

```{r cloze-by-cond-unsmoothed}
dat %>% 
  dplyr::group_by(cond_label) %>% 
  dplyr::summarise(MeanCloze = mean(cloze.target),
                   ClozeSD = sd(cloze.target),
                   MinCloze = min(cloze.target),
                   MaxCloze = max(cloze.target))
```

Stone et al. smoothed:

```{r cloze-by-cond-smoothed}
dat %>% 
  dplyr::group_by(cond_label) %>% 
  dplyr::summarise(MeanCloze = mean(cloze.target_smth),
                   ClozeSD = sd(cloze.target_smth),
                   MinCloze = min(cloze.target_smth),
                   MaxCloze = max(cloze.target_smth))
```

Manually recalculated:

```{r cloze-by-cond-recalculated}
dat %>%
  dplyr::group_by(cond_label) %>%
  dplyr::summarise(MeanCloze = mean(cloze_stemnoun),
                   ClozeSD = sd(cloze_stemnoun),
                   MinCloze = min(cloze_stemnoun),
                   MaxCloze = max(cloze_stemnoun))
```

Visualizations of the manually recalculated stem noun cloze probabilities by condition:

```{r stem-cloze-by-cond-viz}
# Histogram of cloze probs
ggplot(data=dat,
       aes(cloze_stemnoun, fill = cond_label)) + 
  # geom_histogram(bins = 20, fill = "steelblue", color = "steelblue4") + 
  geom_histogram(bins = 20) + 
  scale_fill_manual(values = c("navy", "magenta3")) +
  labs(title = "Histogram of Cloze probabilities by condition") +
  theme_minimal()

# Density plot
ggplot(data=dat,
       aes(cloze_stemnoun, fill = cond_label, color = cond_label)) + 
  geom_density(alpha = 0.6) + 
  scale_fill_manual(values = c("navy", "magenta3")) +
  scale_color_manual(values = c("navy", "magenta3"), guide = "none") +
  labs(title = "Cloze probabilities by condition",
       fill = "Condition") +
  theme_minimal()
```


## Word form cloze

```{r cloze-word-form}
# x <- dat %>% 
#   select(item, cond_label, noun, stemnoun, cloze.target, cloze.target_smth,
#          clozenoun, cloze_stemnoun, wordform_highest, cloze_wordform_highest)

# How often is the cloze probability of the most frequent word form also
# identical to that of the cloze word stem?
summary(round(dat$cloze_wordform_highest, 2) == dat$cloze_stemnoun)
summary(dat$cloze.target == dat$cloze_stemnoun) # stem used by Stone et al. vs. stem used by me

# How often are the actual words identical?
summary(dat$wordform_highest == dat$stemnoun)  # stemnoun: stem used by Stone et al.
summary(dat$wordform_highest == dat$clozenoun) # clozenoun: stem used by me

dat %>%
  dplyr::group_by(cond_label) %>%
  dplyr::summarise(MeanCloze = mean(cloze_wordform_highest),
                   ClozeSD = sd(cloze_wordform_highest),
                   MinCloze = min(cloze_wordform_highest),
                   MaxCloze = max(cloze_wordform_highest))
```

```{r stem-cloze-by-cond-wordform-viz}
# Histogram of cloze probs
ggplot(data=dat,
       aes(cloze_wordform_highest, fill = cond_label)) + 
  # geom_histogram(bins = 20, fill = "steelblue", color = "steelblue4") + 
  geom_histogram(bins = 20) + 
  scale_fill_manual(values = c("navy", "magenta3")) +
  labs(title = "Histogram of Cloze probabilities by condition (word forms)") +
  theme_minimal()

# Density plot
ggplot(data=dat,
       aes(cloze_wordform_highest, fill = cond_label, color = cond_label)) + 
  geom_density(alpha = 0.6) + 
  scale_fill_manual(values = c("navy", "magenta3")) +
  scale_color_manual(values = c("navy", "magenta3"), guide = "none") +
  labs(title = "Cloze probabilities by condition (word forms)",
       fill = "Condition") +
  theme_minimal()
```



Some of the condition a targets have low cloze probability in condition a when considering only the cloze probability of their word form, not that of the actual stem.  
Take a closer look at these items:

```{r inspect-low-wordform-cloze-a}
dat_low_a_wordform_cloze <- dat %>% 
  filter(cloze_wordform_highest < 0.5 & 
           cond_label == "a") %>% 
  arrange(cloze_wordform_highest)
```

This concerns 17 items.  
It remains unproblematic as long as the highest cloze word form is the stem, like "Bügel" in item 190a, which alone accounts for a cloze probability of 0.3158 of responses, while the longer forms "Kleiderbügel" and "Samtkleiderbügel" have additional cloze probabilities of 0.2632 and 0.0526, respectively;  
However, it's more tricky when the highest cloze word form is the longer variant, like "Drogenspürhund" in item 62a, where it might be problematic to simply use the stem "Hund" instead, or "Alarmanlage" in item 108a, where the other attested word forms are "Sicherheitsanlage" and "Sicherungsanlage", while the stem noun "Anlage" was never used by itself.



# Target length and frequency by condition

```{r check-cloze-freq-length}
dat %>% 
  dplyr::group_by(cond_label) %>% 
  dplyr::summarise(MeanLength = mean(target_length),
                   MinLength = min(target_length),
                   MaxLength = max(target_length),
                   SDLength = sd(target_length))

dat %>% 
  dplyr::group_by(cond_label) %>% 
  dplyr::summarise(MeanFreq = mean(as.numeric(ZipfSUBTLEX), na.rm = TRUE),
                   MinFreq = min(as.numeric(ZipfSUBTLEX), na.rm = TRUE),
                   MaxFreq = max(as.numeric(ZipfSUBTLEX), na.rm = TRUE),
                   SDFreq = sd(as.numeric(ZipfSUBTLEX), na.rm = TRUE))
```

A NOTE OF CAUTION: Many frequencies could not be found in the German Subtlex corpus, therefore the above frequency estimate is to be treated with caution.

Missing frequency observations by condition (these might contain the same noun multiple times):

```{r freq-nas}
dat %>%
  dplyr::group_by(cond_label) %>%
  dplyr::summarize(na_count = sum(is.na(ZipfSUBTLEX)))
```


# Duplicate targets

```{r check-duplicate-nouns}
duplicates <- dat %>%
  dplyr::mutate(noun = gsub("[[:punct:]]", "", noun)) %>% 
  dplyr::group_by(noun) %>%
  dplyr::summarize(occurrences = n(),
            items = paste(unique(item), collapse = ", "),
            conditions = paste(unique(cond_label), collapse = ", ")) %>%
  dplyr::filter(occurrences > 1) %>% 
  arrange(-occurrences)

head(duplicates, n = 20)
cat(nrow(duplicates), "nouns are used more than once in the stimuli.")
unique(sort(duplicates$noun))

more_than_twice <- sum(duplicates$occurrences[duplicates$occurrences > 2])
more_than_twice_keep_two <- sum(duplicates$occurrences[duplicates$occurrences > 2] - 2)
cat(more_than_twice, "items contain nouns that are used more than twice.")
cat(more_than_twice_keep_two, "of these should be removed if two occurrences of each target can be kept.")
```

Out of the 158 items kept, 16 must be eliminated to avoid target nouns from appearing more than twice.

**A note of caution:** Note that this check does NOT include checking for reuse of target nouns in the context sentences of other items.

TODO In addition, the 22 items of the discarded items that could potentially be recovered must be checked for target overlap too. It may be that some of these must be discarded due to reuse of target nouns as well.


# Similarity metrics

## Target-context similarity

Check Stone et al. context similarity (LSA):

```{r check-cosine-context}
dat %>% 
  dplyr::group_by(cond_label) %>% 
  dplyr::summarise(MeanSimilarity = mean(cosine_targetContext, na.rm = TRUE),
                   MinSimilarity = min(cosine_targetContext, na.rm = TRUE),
                   MaxSimilarity = max(cosine_targetContext, na.rm = TRUE),
                   SDSimilarity = sd(cosine_targetContext, na.rm = TRUE))
```


## Target-target similarity: embeddings

### Fasttext

Visualize/inspect the fasttext embeddings:

```{r check-embeddings}
fasttext <- dat %>% 
  select(item, cond_label, noun, fasttext) %>% 
  unique() %>% 
  pivot_wider(names_from = cond_label, values_from = noun)


# Histogram of fasttext embeddings
ggplot(data=fasttext,
       aes(fasttext)) + 
  geom_histogram(bins = 20, fill = "steelblue", color = "steelblue4") + 
  labs(title = "Histogram of fasttext target-target similarities") +
  theme_minimal()
```

Manually inspect which target noun pairs exhibit particularly high similarity:

```{r check-high-fasttext}
mean(fasttext$fasttext)
median(fasttext$fasttext)
range(fasttext$fasttext)

fasttext_high <- fasttext %>% 
  filter(fasttext >= 0.5)

nrow(fasttext_high)

# fasttext_high %>% 
#   arrange(-fasttext)
fasttext %>% 
  arrange(-fasttext)
```

Result: 24 items exhibit particularly high fasttext similarity scores of $>=0.5$, among which many pairs corresponding to intuition, such as:

- Maßband - Lineal (0.69)
- Katze - Ratte (0.69)
- Anzug - Sakko (0.67; 2x)

The lowest similarity scores in the item set are observed for:

- Geschirrtuch-Mittel	(0.11)
- Gesangsbuch-Brötchen (0.11)
- Taschenrechner-Schieber (0.13)
- Fächer-Fön (0.13)
- Ladekabel-Teil (0.13)

To filter them out later, a high-fasttext flag is added to those items. 
In addition, the fasttext values are exported sorted by decreasing fasttext similarity score for manual inspection.

```{r flag-high-fasttext-export}
dat$fasttext_high <- ifelse(dat$item %in% fasttext_high$item, 1, 0)
fasttext <- fasttext %>% 
  arrange(-fasttext)
write.csv(fasttext, 'stimuli_fasttext_embeddings_decreasing_keepitems.csv',
          row.names = FALSE)
```

Result of manual inspection: Even among the items with a similarity score of $<=0.5$, many seem critical.  
Manual inspection intuitively suggests to choose a cutoff value around $0.42$ or $0.40$, but even below this threshold value, some items might need to be manually excluded whose similarity could not be captured by fastText efficiently, such as 'Fußbilz-Erreger' ($0.397$) or 'Namensschild-Hemd' ($0.37$).

```{r check-high-fasttext-0.4}
fasttext_high <- fasttext %>% 
  filter(fasttext >= 0.4)
length(unique(fasttext_high$item))
```

Using a threshold of $>=0.40$ for exclusion would however result in the exclusion of 56 items...


### Glove 

Visualize/inspect the GloVe embeddings:

```{r read-glove}
glove <- read.csv("stimuli_glove_embeddings.csv", header = TRUE)

mean(glove$glove, na.rm = TRUE)
median(glove$glove, na.rm = TRUE)
range(glove$glove, na.rm = TRUE)

summary(is.na(glove$glove))
```


```{r check-embeddings-glove}
# Histogram of glove embeddings
ggplot(data=glove,
       aes(glove)) + 
  geom_histogram(bins = 20, fill = "steelblue", color = "steelblue4") + 
  labs(title = "Histogram of glove target-target similarities") +
  theme_minimal()
```

Manually inspect which target noun pairs exhibit particularly high similarity:

```{r check-high-glove}
glove %>% 
  arrange(-glove)

glove_high <- glove %>% 
  filter(glove >= 0.4)
length(unique(glove_high$item))
```


### Comparing GloVe & Fasttext

```{r cor-embeddings}
dat_embeddings <- merge(fasttext, glove) %>% 
  filter(!(is.na(glove)))
cor(dat_embeddings$glove, dat_embeddings$fasttext)
```

Low correlation between the GloVe and fasttext embeddings. The German GloVe embeddings used appear unreliable.



# Surprisal

Define functions to analyze surprisal:

```{r check-surprisal-functions}
correlation_plot <- function(dat, cloze, surprisal) {
  model <- gsub("^surprisal_", "", surprisal)
  p <- ggplot(dat, aes(x = eval(parse(text = cloze)),
                       y = eval(parse(text = surprisal)),
                       color = cond_label)) +
    geom_point(size = 3) + 
    geom_text(aes(label = noun), vjust = -0.5, size = 4) +
    geom_smooth(method = "lm", formula = y ~ x, color = "darkred", se = TRUE) +
    labs(
      title = "Cloze-Surprisal Correlation",
      subtitle = model,
      x = "Target Cloze Probability",
      y = paste0("Target Surprisal (", model, ")"),
      color = "Condition"
    ) +
    scale_color_manual(values =  c("navy", "magenta3")) +
    theme_minimal()
  return(p)
}

get_surprisal_summary <- function(dat, surprisal) {
  dat %>% 
    dplyr::group_by(cond_label) %>% 
    dplyr::summarise(MeanSurprisal = mean(eval(parse(text = surprisal)),
                                          na.rm = TRUE),
                     SDSurprisal = sd(eval(parse(text = surprisal)),
                                      na.rm = TRUE),
                     MinSurprisal = min(eval(parse(text = surprisal)),
                                        na.rm = TRUE),
                     MaxSurprisal = max(eval(parse(text = surprisal)),
                                        na.rm = TRUE))
}

surprisal_density_plot <- function(dat,
                                   surprisal,
                                   plot_title = "Surprisal by condition") {
  model <- gsub("^surprisal_", "", surprisal)
  ggplot(data=dat,
         aes(eval(parse(text = surprisal)),
             fill = cond_label,
             color = cond_label)) + 
    geom_density(alpha = 0.6) + 
    scale_fill_manual(values = c("navy", "magenta3")) +
    scale_color_manual(values = c("navy", "magenta3"), guide = "none") +
    labs(title = plot_title,
         subtitle = model,
         x = paste0("Surprisal (", model, ")"),
         y = "Density",
         fill = "Condition") +
    theme_minimal()
}

surprisal_by_cond <- function(dat, surprisal) {
  dat_surprisal <- dat %>% 
    dplyr::group_by(item, cond_label) %>%
    dplyr::summarise(!!sym(surprisal) := mean(!!sym(surprisal)), .groups = "drop") %>%
    pivot_wider(names_from = cond_label, values_from = !!sym(surprisal))
  
  dat_strong <- dat_surprisal %>%
    filter(a < b)
  dat_strong <- dat %>% 
    filter(item %in% dat_strong$item)
  dat_weak <- dat_surprisal %>%
    filter(a >= b)
  dat_weak <- dat %>% 
    filter(item %in% dat_weak$item)
  
  cat("Checking surprisal by condition for model:",
      surprisal, end = "\n")
  cat("Number of items with surprisal in A < B:",
      length(unique(dat_strong$item)), end = "\n")
  cat("Number of items with surprisal in A >= B:",
      length(unique(dat_weak$item)), end = "\n")
  
  # Check item subset that fulfill the surprisal-condition criterion:
  cat("Mean surprisal by condition (only items with surprisal A <= B):",
      end = "\n")
  print(get_surprisal_summary(dat_strong, surprisal))
  
  # Density plot
  cat("Density plot (only items with surprisal A <= B):", end = "\n")
  surprisal_density_plot(dat_strong, surprisal,
                         plot_title = "Surprisal by condition (good items subset)")
} 
```


## Correlations

### Correlation between model surprisal values

Check correlations between the surprisal values produced by the different models:

```{r surprisal-correlations}
dat_cor <- dat %>%
  rename_with(~ str_remove(., "^surprisal_")) %>% 
  select(gerpt2, gerpt2_large, 
         leo_7b, leo_13b) %>% 
  cor() %>% 
  print()

corrplot(dat_cor, method = "color", type = "lower",
         col = colorRampPalette(c("blue", "white", "red"))(200), 
         tl.cex = 0.8, number.cex = 0.7, addCoef.col = "black", tl.srt = 45)
```

### Correlation between cloze and surprisal

#### Stem nouns (Stone et al. stems)

Check correlations between surprisal and cloze values:

```{r surprisal-cloze-correlations-stemnoun}
dat_cor <- dat %>%
  select(cloze.target,
         surprisal_gerpt2, surprisal_gerpt2_large, 
         surprisal_leo_7b, surprisal_leo_13b) %>% 
  cor() %>% 
  as.data.frame() %>% 
  select(cloze.target) %>% 
  rownames_to_column(var = "model") %>%
  filter(model != "cloze.target") %>%
  mutate(model = str_remove(model, "^surprisal_")) %>% 
  mutate(model = factor(model, 
                        levels = c("gerpt2", "gerpt2_large", 
                                   "leo_7b", "leo_13b"))) %>% 
  print()

ggplot(dat_cor, aes(x = model,
                    y = cloze.target,
                    group = 1)) +
  geom_line(color = "steelblue") + 
  geom_point(color = "steelblue", size = 3) +
  geom_text(aes(label = round(cloze.target, 2)),
                vjust = -1, size = 4) +
  theme_minimal() +
  labs(title = "Surprisal-Cloze Correlations",
       x = "Model",
       y = "Correlation") +
  ylim(-1, 0) +                   # Set y-axis limits
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r plot-surprisal-cloze-correlations-stemnouns, fig.height = 20, fig.width = 20}
p1 <- correlation_plot(dat, "cloze.target", "surprisal_gerpt2")
p2 <- correlation_plot(dat, "cloze.target", "surprisal_gerpt2_large")
p3 <- correlation_plot(dat, "cloze.target", "surprisal_leo_7b")
p4 <- correlation_plot(dat, "cloze.target", "surprisal_leo_13b")
gridExtra::grid.arrange(p1, p2, p3, p4, nrow=2)
```


#### Stem nouns (my stems + corrected cloze values)

Check correlations between surprisal and cloze values:

```{r surprisal-cloze-correlations-mystems}
dat_cor <- dat %>%
  select(cloze_stemnoun,
         surprisal_gerpt2, surprisal_gerpt2_large, 
         surprisal_leo_7b, surprisal_leo_13b) %>% 
  cor() %>% 
  as.data.frame() %>% 
  select(cloze_stemnoun) %>% 
  rownames_to_column(var = "model") %>%
  filter(model != "cloze_stemnoun") %>%
  mutate(model = str_remove(model, "^surprisal_")) %>% 
  mutate(model = factor(model, 
                        levels = c("gerpt2", "gerpt2_large", 
                                   "leo_7b", "leo_13b"))) %>% 
  print()

ggplot(dat_cor, aes(x = model,
                    y = cloze_stemnoun,
                    group = 1)) +
  geom_line(color = "steelblue") + 
  geom_point(color = "steelblue", size = 3) +
  geom_text(aes(label = round(cloze_stemnoun, 2)),
                vjust = -1, size = 4) +
  theme_minimal() +
  labs(title = "Surprisal-Cloze Correlations",
       x = "Model",
       y = "Correlation") +
  ylim(-1, 0) +                   # Set y-axis limits
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r plot-surprisal-cloze-correlations-mystems, fig.height = 20, fig.width = 20}
p1 <- correlation_plot(dat, "cloze_stemnoun", "surprisal_gerpt2")
p2 <- correlation_plot(dat, "cloze_stemnoun", "surprisal_gerpt2_large")
p3 <- correlation_plot(dat, "cloze_stemnoun", "surprisal_leo_7b")
p4 <- correlation_plot(dat, "cloze_stemnoun", "surprisal_leo_13b")
gridExtra::grid.arrange(p1, p2, p3, p4, nrow=2)
```


#### Word forms

```{r surprisal-cloze-correlations-wordforms}
dat_cor <- dat %>%
  select(cloze_wordform_highest,
         surprisal_gerpt2, surprisal_gerpt2_large, 
         surprisal_leo_7b, surprisal_leo_13b) %>% 
  cor() %>% 
  as.data.frame() %>% 
  select(cloze_wordform_highest) %>% 
  rownames_to_column(var = "model") %>%
  filter(model != "cloze_wordform_highest") %>%
  mutate(model = str_remove(model, "^surprisal_")) %>% 
  mutate(model = factor(model, 
                        levels = c("gerpt2", "gerpt2_large", 
                                   "leo_7b", "leo_13b"))) %>% 
  print()

ggplot(dat_cor, aes(x = model,
                    y = cloze_wordform_highest,
                    group = 1)) +
  geom_line(color = "steelblue") + 
  geom_point(color = "steelblue", size = 3) +
  geom_text(aes(label = round(cloze_wordform_highest, 2)),
                vjust = -1, size = 4) +
  theme_minimal() +
  labs(title = "Surprisal-Cloze Correlations",
       x = "Model",
       y = "Correlation") +
  ylim(-1, 0) +                   # Set y-axis limits
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r plot-surprisal-cloze-correlations-wordforms, fig.height = 20, fig.width = 20}
p1 <- correlation_plot(dat, "cloze_wordform_highest", "surprisal_gerpt2")
p2 <- correlation_plot(dat, "cloze_wordform_highest", "surprisal_gerpt2_large")
p3 <- correlation_plot(dat, "cloze_wordform_highest", "surprisal_leo_7b")
p4 <- correlation_plot(dat, "cloze_wordform_highest", "surprisal_leo_13b")
gridExtra::grid.arrange(p1, p2, p3, p4, nrow=2)
```





## Surprisal by condition & model

### GerPT2

```{r gerpt2-surprisal}
get_surprisal_summary(dat, "surprisal_gerpt2")
surprisal_density_plot(dat, "surprisal_gerpt2")
surprisal_by_cond(dat, "surprisal_gerpt2")
```

The full distributions are not very different between conditions.  
Manual inspection reveals that often surprisal is higher in condition a vs. b.
This is the case for 67 of the 158 items.  


### GerPT2-large

```{r gerpt2-large-surprisal}
get_surprisal_summary(dat, "surprisal_gerpt2_large")
surprisal_density_plot(dat, "surprisal_gerpt2_large")
surprisal_by_cond(dat, "surprisal_gerpt2_large")
```

GerPT2-large values lead to an improved alignment of surprisal with cloze values, but there are still 37 out of 158 items for which surprisal in condition A is larger or equal to that in condition B.


### Leo-7B

```{r leo_7b-surprisal}
get_surprisal_summary(dat, "surprisal_leo_7b")
surprisal_density_plot(dat, "surprisal_leo_7b")
surprisal_by_cond(dat, "surprisal_leo_7b")
```

For Leo-7B, only 17 items have higher/equal surprisal in condition A vs. B.

### Leo-13B

```{r leo_13b-surprisal}
get_surprisal_summary(dat, "surprisal_leo_13b")
surprisal_density_plot(dat, "surprisal_leo_13b")
surprisal_by_cond(dat, "surprisal_leo_13b")
```

For Leo-13B, there are also 17 items with surprisal A >= B.


### Check all models

Check the items for which the strongest model (leo13b) yielded surprisal b <= a:

```{r check-bad-surprisal-items}
dat_surprisal_leo <- dat %>% 
  dplyr::group_by(item, cond_label) %>%
  dplyr::summarise("surprisal_leo_13b" := mean(surprisal_leo_13b), .groups = "drop") %>%
  pivot_wider(names_from = cond_label, values_from = surprisal_leo_13b)

dat_weak_leo <- dat_surprisal_leo %>%
  filter(a >= b)
dat_weak_leo <- dat %>% 
  filter(item %in% dat_weak_leo$item) %>% 
  select(item, cond_label, sentence_target, noun,
         target_length, ZipfSUBTLEX,
          cloze_stemnoun,
         surprisal_gerpt2, surprisal_gerpt2_large,
         surprisal_leo_7b, surprisal_leo_13b) %>% 
  arrange(item, cond_label)
```

Check the items for which at least 3 out of 4 models yielded surprisal b <= a:

```{r check-bad-surprisal-all-models}
dat_surprisal <- dat %>% 
  dplyr::group_by(item, cond_label) %>%
  dplyr::summarise("surprisal_gerpt2" := mean(surprisal_gerpt2),
                   "surprisal_gerpt2_large" := mean(surprisal_gerpt2_large), 
                   "surprisal_leo_7b" := mean(surprisal_leo_7b),
                   "surprisal_leo_13b" := mean(surprisal_leo_13b), .groups = "drop") %>%
  pivot_wider(names_from = cond_label, values_from = c(surprisal_gerpt2, surprisal_gerpt2_large,
                                                       surprisal_leo_7b, surprisal_leo_13b))

dat_weak <- dat_surprisal %>%
  filter(rowSums(cbind(surprisal_gerpt2_a >= surprisal_gerpt2_b,
                       surprisal_gerpt2_large_a >= surprisal_gerpt2_large_b,
                       surprisal_leo_7b_a >= surprisal_leo_7b_b,
                       surprisal_leo_13b_a >= surprisal_leo_13b_b)) >= 3)

dat_weak <- dat %>% 
  filter(item %in% dat_weak$item) %>% 
  select(item, cond_label, sentence_target, noun,
         target_length, ZipfSUBTLEX,
         cloze_stemnoun,
         surprisal_gerpt2, surprisal_gerpt2_large,
         surprisal_leo_7b, surprisal_leo_13b) %>% 
  arrange(item, cond_label)
```

Based on manual inspection of these items, it does not seem necessary to exlude items on the basis of surprisal.  
The surprisal values do not seem to point at particularly problematic items.